---
title: "JAGS example"
author: "Paul M"
date: "3/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to JAGS  ("Just Another GIBBS Sampler")

[Based on one of the examples at http://www.johnmyleswhite.com/notebook/2010/08/20/using-jags-in-r-with-the-rjags-package/]

First you must install jags on your computer! rjags won't run without it.
e.g., in the terminal using homebrew: brew install jags"
Then install the rjags package:
```{r}
#install.packages("rjags")
library(rjags)
library(coda)
```


# An example using normal rvs, where we are infering the mean and variance:

Let's asusme we have a set of rvs that we know are normally distributed, but wer wish to infer their mean and variance. Generate a test set of data
```{r}
N <- 1000
x <- rnorm(N, 0, 2)
 
write.table(x,
            file = 'example1.data',
            row.names = FALSE,
            col.names = FALSE)
```

Now we need to write a model specification in JAGS syntax. Put the model specification (below) in a file called example1.bug. The complete model looks like this:

model {
	for (i in 1:N) {  
           x[i] ~ dnorm(mu, tau)  
    	}  
    	mu ~ dnorm(0, .0001)  
	    tau <- pow(sigma, -2)  
	    sigma ~ dunif(0, 100)  
}

* The first line says that you are specifying a model. 
* Then you set up the model for every single data point using a for loop. 
Here, we say that x[i] is distributed normally with mean mu and precision tau.
Note that rjags works with precision rather than variance! 
(Precision=1/variance)
*Then we specify priors for mu and tau. 
* mu is assumed to be distributed normally with mean 0 and standard deviation 100. This is an example of a  non-informative prior. 
+ Then we specify tau as a deterministic function (hence the deterministic <- instead of the distributional ~) of sigma, after raising sigma to the -2 power. Then we say that sigma has a uniform prior over the interval [0,100].

Now we invoke the model:
```{r}
# Set up our model object in R
jags <- jags.model('example1.bug',   # specification file
                   data = list('x' = x,   # the data (must use same names as trhe model spec. file)
                               'N' = N),
                   n.chains = 4, # how many parallel chains to run
                   n.adapt = 100   # it can use adaptive sampling, removing the first 100 iterations
                   )
 
update(jags, 1000)    # run another 1000 iterations. The update function is used to add more iterations.
 
samps <- jags.samples(jags,   # draw 1000 samples from the output for the requested variables
             c('mu', 'tau'),
             1000)

summary(samps)
```


You can also set it for use of coda diagnostics for covergence (which also seems to provide a better summary of the output)
```{r}
samps2 <- coda.samples( jags, c('mu','tau'), 1000 )
summary(samps2)
plot(samps2)
show(gelman.diag(samps2))

```
